Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-19 csf2/main.tex
some bnfs

\bul{A cross-language relation $\sim$ is well-formed with respect to a class of source-level properties $\src{\class}$} iff \rul{the existential image preserves set membership}.
\begin{definition}[Well-formedness of Trace Relations for a Class $\src{\class}$]\label{def:wfc:tau:tracerel}
  % $$

  \noindent
  \text{\bul{$\wfctau{\sim}{\src{\class}}$}} \isdef \text{\rul{$\forall \src{\Pi}\in\src{\class}, \tau_\sim(\src{\Pi})\in\tilde{\tau}_\sim(\src{\class})$}}
  % $$
\end{definition}

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\tau$]\label{thm:rtpsim:tau}
  If \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtptau{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wfctau{\sim_1}{\src{\class[_2]}}$}, then \oul{$\rtptau{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

The dual theorem also holds: let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

When the cross-language trace relation is an equality, \Cref{thm:rtpsim:tau,thm:rtpsim:sig} collapse:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

\begin{corollary}[Sequential Composition of Secure Compilers]\label{corr:swappable}
  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}. \Coqed
\end{corollary}

@@

This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later.

\subsection{Basic Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  % \footnotesize
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
                              \ev{Use\ \loc\ n;\sandboxtag}&\le_{\trace}&\ev{Dealloc\ \loc;\sandboxtag} \\
                            %  \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc;\sandboxtag}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n;\sandboxtag}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

  \noindent
  $$
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag}, \text{ then }m<n
                                \end{array}\right.\right\}
                                $$
  % \end{nscenter}
\end{definition}

@@

\subsection{Memory Safety Trace Model}

@@

  \noindent$$
  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                 \end{array}\right.\right\}
  $$
% \end{nscenter}
\end{definition}

@@

The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
\[
  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sandboxtag\sim_{msscct}\event[_b];\sandboxtag;\securitytag}
\]
\[
  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\sandboxtag;\securitytag}
  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\sandboxtag;\securitytag}
\]

@@

\subsection{Speculation Trace Model}

@@

\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
  \noindent\[
  \sssafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\sandboxtag;\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
                                 \end{array}\right.\right\}
  \]
\end{definition}

@@

Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
To this end, let $\sim_{spec}$ be a cross-language trace relation that relates the non-speculative memory safety trace model with the one that has speculation by simply ignoring any events related to speculation and ignoring the additional variant in the security-tag:
\[
  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\sandboxtag;\securitytag}
  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\sandboxtag;\securitytag}
\]
\[
  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\sandboxtag;\lock\sim_{spec}\event[_b];\sandboxtag;\lock_{vX}}
\]

@@

%   \typerule{tms-refl}{
%     \tmsmonitor.\text{allocated}=\emptyset
%   }{
%     \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
%   }{tms-refl}
%   %
%   \typerule{tms-ign-trans}{
%     \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
%   }{
%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
%   }{tms-ign-trans}
%   %
%   \typerule{tms-trans}{
%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent} &
%     \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
%   }{
%     \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
%   }{tms-trans}
% \end{center}
% \Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that all locations should be deallocated.
% This rejects the behavior of programs that forget to free memory.

% \begin{center}\small
%   \judgbox{\tmstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ w.r.t. \gls*{tms}.''}$\;$\\
%   %
%   \typerule{tms-alloc-authentic}{
%   }{
%     \tmstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ \sandboxtag}
%   }{tms-alloc-authentic}
%   %
% % \typerule{tms-dealloc-authentic}{
% % }{
% %   \tmstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
% % }{tms-dealloc-authentic}
%   %
% % \typerule{tms-use-authentic}{
% % }{
% %   \tmstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc}
% % }{tms-use-authentic}
%   %
%   \typerule{tms-branch-authentic}{
%   }{
%     \tmstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{tms-branch-authentic}
%   %
% % \typerule{tms-binop-authentic}{
% % }{
% %   \tmstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
% % }{tms-binop-authentic}
% % %
% % \typerule{tms-empty-authentic}{
% % }{
% %   \tmstraceagree[]{\emptyevent}{\absev{\emptyevent}}
% % }{tms-empty-authentic}
% % %
%   % \typerule{tms-abort-authentic}{
%   % }{
%   %   \tmstraceagree[]{\lightning}{\absterm}
%   % }{tms-abort-authentic}
% \end{center}

% The trace agreement maps actions to their respective (e.g., \Cref{tr:tms-alloc-authentic}), so we elide most cases.
% % 
% Since branches are not relevant for \gls*{tms}, they are mapped to $\absev{\emptyevent}$ (\Cref{tr:tms-branch-authentic}).
% % \MP{
% % 	we should say something .. ?
% % }

% \begin{lemma}[Monitor Satisfaction is $\tmssafe$]\label{lem:mon:tmssafe}
%   $\;\;$ If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
% \end{lemma}

% \begin{exampleenv}[A program not satisfying \gls*{tms}]\label{ex:strncpy:cpp11}
%   Consider the following \texttt{C++11} library that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.

%   \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
% int hello() {
%   char* greet = new char[12] { "Hello ESOP!" };
%   char* to = new char[12];
%   strncpy(11, to, greetings);
%   delete to;
%   printf("%cSOP\n", to[6]);
% }
%   \end{lstlisting}
%     % // greet address ${\color{gray}\loc_{x}}$
%   % // to address ${\color{gray}\loc_{y}}$
% % 
%   Up to the body of \texttt{printf}, the program execution yields the specification trace: 
%     % $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$

%     \vspace{-1em}
%     {\small
%     \begin{align*}
%       &
%       \ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock
%       \cdot
%       \ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
%       \\
%       &
%       \cdot
%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
%       \cdot
%       \ldots
%       \\
%       &
%       \cdot
%       \ev{Use}\ \loc_{x}\ 12;\comp;\unlock
%       \cdot
%       \ev{Dealloc}\ \loc[_{y}];\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{y}\ 6;\comp;\unlock
%     \end{align*}
%     }
% Relating this trace to abstract monitor events yields trace:
%   
%   \vspace{-1em}
%   {\small
%   \begin{align*}
%     % \abstrace=
%     &
%     \absev{Alloc}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Alloc}\ \loc[_{y}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{x}]\ \comp
%     % \\
%     % &
%     \cdot
%     \absev{Use}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{y}]\ \comp
%     \cdot
%     \ldots
%     \\
%     &
%     \cdot
%     \absev{Use}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Dealloc}\ \loc[_{y}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{y}]\ \comp
%    \end{align*} 
%     }
%   Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), observe that it does not deallocate its arguments.
%   Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls*{tms}, since location $\loc[_{x}]$ is still allocated.
%   Having returned from \texttt{strncpy}, the \texttt{greet} function continues and deallocates $\loc[_{y}]$ whose subsequent use in \texttt{printf} is a use-after-free bug.

%   The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to trace: 
%   
% \vspace{-1em}
%   {\small
%   \begin{align*}
%     \abstrace[']=
%     &
%     \absev{Alloc}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Alloc}\ \loc[_{y}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{x}]\ \comp
%     % \\
%     % &
%     \cdot
%     \absev{Use}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{y}]\ \comp
%     \cdot
%     \ldots
%     \cdot
%     \\
%     &
%     \absev{Use}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Dealloc}\ \loc[_{x}]\ \comp
%     \cdot
%     \absev{Use}\ \loc[_{y}]\ \comp	
%   \end{align*}
%   }
%   It follows that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls*{sms}.
% \end{exampleenv}

% % \MPpost{
% % 	i think the trace description of traces is more uindertandable,
% % 	teh monitor one is easier to work with
% % 	this is the justification for having both that we'll have to make in the future
% % }

% \subsubsection{Monitor for \glsfirst*{sms}}\label{subsubsec:sms}
% \begin{gather*}
%   \begin{aligned}
%     \mi{(Abstract\ Store)}~\smsmonitor:=&\ L\times\sandboxtag\times\mb{N}\\
%     \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
%   \end{aligned}
% \end{gather*}
% \begin{center}
%   \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ steps to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
%   %
% % \typerule{sms-uninteresting}{
% % }{
% %   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% % }{sms-uninteresting}
% % %
% % \typerule{sms-abort}{
% % }{
% %   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% % }{sms-abort}
%   %
%   \typerule{sms-use}{
%     (\loc;\sandboxtag;m)\in\smsmonitor &
%     n<m
%   }{
%     \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
%   }{sms-use}
%   %
%   \typerule{sms-alloc}{
%     (\loc;\sandboxtag;m)\notin\smsmonitor %\\
%     % unfortunately, this does not render sms-use and sms-alloc side by side
%     % my only other idea is \scalebox...ugh
%     %\smsmonitor[']=\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}
%   }{
%     \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
%   }{sms-alloc}
% \end{center}
% The state of the monitor for \gls*{sms} is a set containing tuples of locations, control-tags, and the allocation size.
% In comparison to the trace model of the \gls*{tms} monitor, the trace model here is extended by sizing and positional information.
% \Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
% The trace agreement is entirely straightforward and similar to the one for \gls*{tms}.
% % \begin{center}
% %   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls*{sms}.''}$\;$\\
% %   %
% %   \typerule{sms-alloc-authentic}{
% %   }{
% %     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
% %   }{sms-alloc-authentic}
% %   %
% %   \typerule{sms-dealloc-authentic}{
% %   }{
% %     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
% %   }{sms-dealloc-authentic}
% %   %
% %   \typerule{sms-use-authentic}{
% %   }{
% %     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
% %   }{sms-use-authentic}
% %   %
% %   \typerule{sms-branch-authentic}{
% %   }{
% %     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
% %   }{sms-branch-authentic}
% %   %
% %   \typerule{sms-binop-authentic}{
% %   }{
% %     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
% %   }{sms-binop-authentic}
% %   %
% %   \typerule{sms-empty-authentic}{
% %   }{
% %     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
% %   }{sms-empty-authentic}
% %   %
% %   \typerule{sms-abort-authentic}{
% %   }{
% %     \smstraceagree[]{\lightning}{\absterm}
% %   }{sms-abort-authentic}
% % \end{center}

% \begin{lemma}[Monitor Satisfaction is $\smssafe$]\label{lem:mon:smsafe}
%   $\;\;$
%   If $\monsmssafe{\trace}$, then $\trace\in\smssafe$. %\Coqed
% \end{lemma}

% \begin{exampleenv}[Normal invocation of \texttt{strncpy}]\label{ex:strncpy:sms}
%   Consider the insecure \texttt{strncpy} function from \Cref{ex:strncpy} with a context \texttt{strncpy(2, x, y)}, where \texttt{x} and \texttt{y} are pointers to valid regions of memory with allocated space for exactly two cells and do not contain the null-terminating character \texttt{'\textbackslash 0'}.
%   For the sake of this example, the pointers have been allocated by the component and passed to the context.
%   The loop of \texttt{strncpy} will copy exactly two cells and then check the loop condition for the last time.
%   At that stage, variable \texttt{i} is equal to \texttt{2} and, unfortunately, the order of checks is such that first the cell \texttt{x[i]} is read prior to bounds checking \texttt{i < n}, so there is an out-of-bounds memory access right before exiting the function.
%   This is also visible on the trace:
%     
%     {\small
%     \begin{align*}
%     &
%       \ldots
%       \cdot
%       \ev{Alloc}\ \loc_{x}\ 2;\comp;\unlock
%       \cdot
%       \ldots
%       \cdot
%       \ev{Alloc}\ \loc_{y}\ 2;\comp;\unlock
%       \cdot
%       \ldots
%       \cdot
%     % \\
%     % &
%       \ev{Use}\ \loc_{x}\ 0;\comp;\unlock
%       \cdot
%       \\
%     &
%       \ev{Use}\ \loc_{y}\ 0;\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{x}\ 1;\comp;\unlock
%     % \\
%     % &
%       \cdot
%       \ev{Use}\ \loc_{y}\ 1;\comp;\unlock
%       \cdot
%       \ev{Use}\ \loc_{x}\ 2;\comp;\unlock
%       \cdot
%       \ldots	
%     \end{align*}
%     }

% \noindent
% where $\loc_{x}$ and $\loc_{y}$ are the memory addresses associated to \texttt{x} and \texttt{y}, respectively.
%   Omitting the events for all ,,$\ldots$'' for sake of brevity, the abstract monitor trace of this is:
%     % 
%     {\small
%     \begin{align*}
%       &
%       \absev{Alloc}\ \loc_{x}\ \comp\ 2
%       \cdot
%       \absev{Alloc}\ \loc_{y}\ \comp\ 2
%       \cdot
%       \absev{Use}\ \loc_{x}\ \comp\ 0
%       \\
%       &
%       \cdot
%       \absev{Use}\ \loc_{y}\ \comp\ 0
%       \cdot
%       \absev{Use}\ \loc_{x}\ \comp\ 1
%       \cdot
%       \absev{Use}\ \loc_{y}\ \comp\ 1
%       \cdot
%       \absev{Use}\ \loc_{x}\ \comp\ 2
%     \end{align*}
%     }

%   After the allocation events, the state of the monitor is $\left\{(\loc_{x};\comp;2),(\loc_{y};\comp;2)\right\}$.
%   All uses up to the last are accepted by the monitor, but the last event does not satisfy the premise $2<2$ in \Cref{tr:sms-use}.
%   Therefore, the whole program (\texttt{strncpy} linked with this kind of context) is not \gls*{sms}.
% \end{exampleenv}

% \subsubsection{Combining \glsfirst*{tms} and \glsfirst*{sms} Monitors into \glsfirst*{ms}}\label{subsubsec:ms}

>>>>

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed with respect to class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers with respect to $\sigma$]\label{thm:rtpsim:sig}
  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_1}{\trg{\class[_2]}}$}, then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

When the cross-language trace relation is an equality, \Cref{thm:rtpsim:sig} collapse:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

\begin{corollary}[Swapping Secure Compiler Passes]\label{corr:swappable}
  If \bul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{1}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc[_{1}]{\trg{L}}{\trg{L}}\circ\cc[_{2}]{\trg{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$ and $\rtp{\cc[_{2}]{\trg{L}}{\trg{L}}\circ\cc[_{1}]{\trg{L}}{\trg{L}}}{\class[_{2}]\cap\class[_{1}]}$}. \Coqed
\end{corollary}

@@

This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) presented later. 
For each of these properties, the technical appendix accompanying the paper defines corresponding monitors, which refine a given property.

\subsection{Memory Safety Trace Model}\label{subsec:propdefs}\label{subsec:basic:memsafety:tracemodel}

\vspace{-1.5em}

\[
  \begin{array}{rcll}
    \event[_{ms}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]

@@

\begin{definition}[\glsfirst*{tms}]\label{def:trace:tmsdef}
  % \footnotesize
  $$
  \tmssafe:=\left\{\trace \left| \begin{array}{rcl}
                              \ev{Alloc\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
                              \ev{Use\ \loc\ n}&\le_{\trace}&\ev{Dealloc\ \loc} \\
                            %  \text{if }\ev{Alloc\ \loc\ n;\sandboxtag}\text{ in }\trace&\text{then}&\ev{Dealloc\ \loc;\sandboxtag}\text{ in }\trace \\
                              \text{at most one }\ev{Dealloc\ \loc}&\text{in}&\trace \\
                              \text{at most one }\ev{Alloc\ \loc\ n}&\text{in}&\trace \\
                            \end{array}\right.\right\}
  $$
\end{definition}
Hereby, the notation $\event[_{1}]\le_{\trace}\event[_{2}]$ means that if $\event[_{1}]$ is in $\trace$ and if $\event[_{2}]$ is in $\trace$, then $\event[_{1}]$ appears before $\event[_{2}]$.

@@

  \noindent
  $$
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n}\le_{\trace}\ev{Use\ \loc\ m}, \text{ then }m<n
                                \end{array}\right.\right\}
                                $$
  % \end{nscenter}
\end{definition}

@@

\subsection{Memory Safety Trace Model with Constant Timeness}

\[
  \begin{array}{rcll}
    \event[^b_{ct}] &:=& \\
    \event[_{ct}] &:=& \ev{Alloc \loc\ n} \mid \ev{Dealloc \loc} \mid \ev{Use \loc\ n} \mid \emptyevent \mid \lightning \\ 
  \end{array}
\]

@@

  \noindent$$
  \scctsafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\scctsafe
                                 \end{array}\right.\right\}
  $$
% \end{nscenter}
\end{definition}

@@

The combination of \gls*{ms} and \gls*{scct} is, intuitively, the intersection of these properties, \gls*{msscct}.
However, \gls*{ms} uses a simpler trace model than \gls*{scct}, so intersecting them would trivially yield the empty set. 
To remedy this issue, let $\sim_{msscct}$ be a cross-language trace relation that relates the basic memory safety trace model with the memory safety trace model by ignoring security tags and replacing the new pre-events with the empty event:
\[
  \typerulenolabel{mscctrel:drop:tag}{}{\event[_b];\sim_{msscct}\event[_b];\securitytag}
\]
\[
  \typerulenolabel{mscctrel:drop:branch}{}{\emptyevent\sim_{msscct}\ev{Branch n};\securitytag}
  \typerulenolabel{mscctrel:drop:binop}{}{\emptyevent\sim_{msscct}\ev{Binop n};\securitytag}
\]

@@

\subsection{Memory-Safe, Constant-Time Trace Model with Speculation}

@@

\begin{definition}[\glsfirst*{ss}]\label{def:trace:ss}
  \noindent\[
  \sssafe:=\left\{\trace \left|\begin{array}{rcl}
                                   \trace=\hole{\cdot} & \text{ or } & \trace=\event[_{b}];\unlock\cdot\trace['] \wedge \trace[']\in\sssafe
                                 \end{array}\right.\right\}
  \]
\end{definition}

@@

Similar to \gls*{msscct}, there is a need to relate the different trace models with each other, so that the memory safety property without speculation can be lifted to speculation. 
To this end, let $\sim_{spec}$ be a cross-language trace relation that relates the non-speculative memory safety trace model with the one that has speculation by simply ignoring any events related to speculation and ignoring the additional variant in the security-tag:
\[
  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{spec}\ev{Spec};\securitytag}
  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{spec}\ev{Rlb};\securitytag}
\]
\[
  \typerulenolabel{specrel:drop:tag}{}{\event[_b];\lock\sim_{spec}\event[_b];\lock_{vX}}
\]

@@
